---
title: "Scaling functions"
author: "Ander Alonso"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
vignette: >
  %\VignetteIndexEntry{Scaling functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
library(SME)
library(knitr)
```

```{r}
df <- data.frame("V1" = 1:3, "V2" = 3:5, "V3" = 5:7)
v <- c(1, 3, 5)
m <- matrix(c(1,2,3,3,4,5,5,6,7), ncol = 3, byrow = FALSE)
```

# normalize
The goal of normalization is to transform features to be on a similar scale, implemented method is max-min normalization:

$x_{normalize} = \frac{x - min(x)}{max(x)-min(x)}$

Currently implemented for numeric vectors, numeric matrices and data.frame normalization.

### Usage
normalize(x, margin = 1L)

### Arguments
- x numeric, matrix or dataframe input vector.
- sep Field separator. Default ;
- margin 1 - normalize by rows and 2 - normalize by columns. Default is 1.

### Value
- A normalized numeric or data.frame.

### Examples 

```{r}
s1 <- normalize(df, margin = 1L)
kable(s1)
```


```{r}
s2 <- normalize(df, margin = 2L)
kable(s2)
```


```{r}
s3 <- normalize(m, margin = 1L)
kable(s3)
```


```{r}
s4 <- normalize(m, margin = 2L)
kable(s4)
```

```{r}
s5 <- normalize(v)
kable(s5)
```



# standardize

Standardization rescales data to have a mean (μ) of 0 and standard deviation (σ) of 1 (unit variance).

$x_{standardize} = \frac{x-mean(x)}{\sigma (x)}$

Currently implemented for numeric vectors, numeric matrices and data.frame standardization.

### Usage
standardize(x, margin = 1L)

### Arguments
- x numeric, matrix or dataframe input vector.
- margin 1 - standardize by rows and 2 - standardize by columns. Default is 1

### Value
A standardize numeric or data.frame.

### Examples 

```{r}
s1 <- standardize(df, margin = 1L)
kable(s1)
```


```{r}
s2 <- standardize(df, margin = 2L)
kable(s2)
```


```{r}
s3 <- standardize(m, margin = 1L)
kable(s3)
```


```{r}
s4 <- standardize(m, margin = 2L)
kable(s4)
```

```{r}
s5 <- standardize(v)
kable(s5)
```



