---
title: "Discretize functions"
author: "Ander Alonso"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
vignette: >
  %\VignetteIndexEntry{Discretize functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r}
library(SME)
library(knitr)
```

# discretize

Discretization is the process of transforming a continuous-valued variable into a discrete one by creating a set of contiguous intervals (or equivalently a set of cutpoints) that spans the range of the variable's values. 

There are diferent discretization methods, for example:

- Equal interval width discretization: all bins have equal width.
- Equal frequency discretization: all bins contain the same number of numerical values.
- k-means discretization:  consists of applying the classical K-means clustering to one-dimensional continuous data, k is equal to the number of bins.

### Usage
discretize(x, method = "frequency", num.bins)

### Arguments
- x a numeric vector, matrix or dataframe (continuous variables)
- method discretization method. Available are: "interval" (equal interval width), "frequency" (equal frequency), "cluster" (k-means clustering).
- num.bins number of intervals.


### Value
A discretized vector or dataframe

### Examples
```{r}
v.example.1 <- c(11.5, 10.2, 1.2, 0.5, 5.3, 20.5, 8.4)
v.example.2 <- c(0,4,12,16,16,18,24,26,28)
df.example.3 <- data.frame(c(0,4,12), c(16,16,18), c(24,26,28))
matrix.example.4 <- matrix(c(0,4,12,16,16,18,24,26,28), 3, 3, byrow = FALSE )
```

```{r}
ew.discretize.example1 <- discretize(v.example.1, method="interval", 4)
print(ew.discretize.example1)

ew.discretize.example2 <- discretize(v.example.2, method="interval", 3)
print(ew.discretize.example2)

ew.discretize.example3 <- discretize(df.example.3, method="interval", 3)
kable(ew.discretize.example3)

ew.discretize.example4 <- discretize(matrix.example.4, method="interval", 2)
kable(ew.discretize.example4)

ef.discretize.example1 <- discretize(v.example.1, method="frequency", 4)
print(ef.discretize.example1)

ef.discretize.example2 <- discretize(v.example.2, method="frequency", 3)
print(ef.discretize.example2)

ef.discretize.example3 <- discretize(df.example.3, method="frequency", 3)
kable(ef.discretize.example3)

ef.discretize.example4 <- discretize(matrix.example.4, method="frequency", 2)
kable(ef.discretize.example4)


clustering.discretize.example1 <- discretize(v.example.1, method="clustering", 4)
print(clustering.discretize.example1)

clustering.discretize.example2 <- discretize(v.example.2, method="clustering", 3)
print(clustering.discretize.example2)

clustering.discretize.example3 <- discretize(df.example.3, method="clustering", 3) 
kable(clustering.discretize.example3)

clustering.discretize.example4 <- discretize(matrix.example.4, method="clustering", 2)   
kable(clustering.discretize.example4)
```

