---
title: "Input Output functions"
author: "Ander Alonso"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
vignette: >
  %\VignetteIndexEntry{Input Output functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
library(SME)
library(knitr)
```

# read.df
Read a csv file

### Arguments
- file.path Name of input file
- sep Field separator. Default ;
- max.n.factor maximum number of column characters to be a factor. Default 3.
- header a logical value indicating whether the file contains the names of the variables as its first line. Default FALSE.
- row.names a vector of row names. Default NULL.
- col.names a vector of col names. Default NULL.
- n.rows integer: the maximum number of rows to read in. In case of Header = TRUE, it does not count for the number of read lines.
- transpose If TRUE, transpose the result. Default FALSE.

### Value
- data frame containing a representation of the data in the file.

# write.df

Write a csv file

### Arguments
- df: the object to be written, preferably a matrix or data frame.
- file.path: Name of output file.
- sep: the field separator string. Values within each row of x are separated by this string. Default ";" .
- append: If TRUE, the output is appended to the file. If FALSE, any existing file of the name is destroyed. Defatul FALSE.
- header: if TRUE, name of the columns are written in the first line. Default FALSE.


# Examples
```{r}
df = data.frame("V1" = 1:5, "V2" = c("a", "b", "c", "d", "e"), "V3" = c("alto", "bajo", "bajo", "alto", "alto"))
```


```{r}
write.df(df, "bench1.txt")
r <- read.df(file.path = "bench1.txt")
kable(r)
```

```{r}
write.df(df, "bench2.txt", header = TRUE)
r <- read.df(file.path = "bench2.txt", header = TRUE)
kable(r)
```


```{r}
write.df(df, "bench2.txt", append = TRUE)
r <- read.df(file.path = "bench2.txt", header = FALSE)
kable(r)
```

```{r}
write.df(df, "bench2.txt", sep = ";", header = FALSE, append = TRUE)
r <- read.df(file.path = "bench2.txt", sep = ";", header = TRUE,  max.n.factor = 3, row.names = c('X1', 'X2', 'X3', 'X4', 'X5'), col.names = c('A', 'B', 'C'), n.rows = 5, transpose = FALSE)
kable(r)
```

